Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-05-31 schemes.tex
multicore information added

\label{sec:schemes}

@@

Before the coding density can be estimated it is essential that the encoder
has a good estimate of the decoders \ac{DOF}, but also the remainder of the
total budget. The feedback scheme \ref{eq:feedback_dofs} roughly splits the
remaining \ac{DOF} in halves. This is also the case for the budget that is
split equally among the two new regions. The very last region will be asigned
the full remainder of the total budget.

@@

Some algorithms implement the decoding combining the matrix inversion and the
matrix multiplication, e.g., performing the Gauss-Jordan algorithm over the
coding matrix $\textbf{V}$ while performing, in parallel, row operations on the
coded data $\textbf{C}$. The parallelization of such operations provide speeds
up that are limited for small block sizes ($\leq 2048$ bytes). The reason is
that to operate in a parallel fashion over the same coded packet $C_i$ requires
strained synchronization.

>>>>

%!TEX root = raspi_journal.tex
\label{sec:schemes}

@@

Before the coding density can be estimated it is essential that the encoder
has a good estimate of the decoders \ac{DOF}, but also the remainder of the
total budget. The feedback scheme \ref{eq:feedback_dofs} roughly splits the
remaining \ac{DOF} in halves. This is also the case for the budget that is
split equally among the two new regions. The very last region will be assigned
the full remainder of the total budget.

@@

When designing multithreaded algorithms for network coding operations, it is
possible to implement the decoding by combining the matrix inversion and the
matrix multiplication, e.g., performing the Gauss-Jordan algorithm over the
coding matrix $\textbf{V}$ while performing, in parallel, row operations on the
coded data $\textbf{C}$. However, the parallelization of such operations provide
speeds up that are limited for small block sizes ($\leq 2048$ bytes). The reason
is that to operate in a parallel fashion over the same coded packet $C_i$
requires strained synchronization. In~\cite{5061951} and~\cite{4262451} the
parallelization of the row operations are optimized for \ac{GPU} and \ac{SMP}
systems respectively.

Therefore, to overcome the constraints of a tight synchronization, it is
preferable to explicitly invert the matrix $\textbf{V}$, and then take advantage
of optimizations for matrix multiplications. The authors implemented
in~\cite{wunderlich2015network} an algorithm that adopts the ideas of efficient
\ac{BLAS} operations~\cite{lawson1979basic} reimplementing them for finite
fields operations. The implemented algorithm tries to be cache efficient
maximizing the number of operations performed over a block of fetched data. This
is done by operating on square sub-blocks of the matrices, which exploits the
spatial locality of the data, at least for $\mathcal{O}(n^3)$
algorithms~\cite{golub2012matrix}. The optimal size of the block is
architecture-dependent. The ideal block has a number of operands that fit into
L1 cache, and it is a multiple of the \ac{SIMD} operations size.

The main idea of the implemented algorithm is to represent each one of the sub-
block matrix operations as a base, \textit{kernel}, operation that can be
optimized, e.g., using \ac{SIMD} operations. Each of these kernels, at the same
time, can be represented as a task with inputs and outputs in memory. If the
algorithm is formulated conventionally, and the data operations on the sub-
blocks are reordered, then it is possible to resolve data dependencies with a
\ac{DAG}. For the algorithm, a dependency means, for example, that a sub-block
of the matrix in memory cannot be read unless a previous writing operation was
performed there. When the \ac{DAG} is determined, then, a scheduler assigns to
the available cores the individual kernel operations that have their
dependencies fulfilled. The benefit of this algorithm is that the
synchronization relies only on data dependencies, and it does not requires the
insertion of artificial synchronization points. Using this technique, the matrix
inversion is performed using an algorithm based on LU
factorization~\cite{Dongarra:2011:HPM:2132876.2132885}, and the matrix
multiplication is performed by making the various matrix-matrix multiplications
of the sub-blocks.
